---
title: "Catalogue of Beluga and Narwhal Click Spectra and Spectrograms from SoundTrap Recordings"
subtitle: "Acoustic Event Data Summary"
author: "Marie J. Zahn, Michael Ladegaard, Malene Simon, Kathleen M. Stafford, Taiki Sakai, Kristin L. Laidre"
date: "Date Produced: `r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: 
    latex_engine: pdflatex
    toc: true
    toc_depth: 3
documentclass: article
geometry:
- top=0.75in
- bottom=0.75in
- left=0.5in
- right=0.5in
editor_options: 
  chunk_output_type: console
---

***

**Summary**

This supplementary file provides an overview of the data included in all beluga and narwhal acoustic events (each 1 hour) from SoundTrap recordings collected at the Kong Oscar and Fisher Islands mooring sites. A summary table is provided for each category showing the number of detections for each acoustic event and the sound pressure level (SPL) difference in decibels (dB) between selected third-octave frequency bands.

There are two figures for each event:

- Title: Acoustic event number that provides a unique numeric ID for each 6-hour acoustic event.
- Left: Concatenated click spectrogram where darker orange colors represent higher energy content.
- Right: Mean power spectrum for all detections within the event.

A 512 point FFT and Hanning window were used to produce all figures.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load packages, results = FALSE, include=FALSE}
## load required packages
library(here)
library(ggplot2)
library(PAMpal)
library(dplyr)
```

```{r load data, results = FALSE, include=FALSE}
## load Fisher data
load(here('R/Rdata/1hr/Fish_20kHzPG_0kHzPP_1hr_wavTime_filtered_labeled.rdata'))

## load Kong Oscar data
load(here('R/Rdata/1hr/Kong_20kHzPG_0kHzPP_1hr_wavTime_filtered_labeled.rdata'))

```

```{r update Binaries location on path, results = FALSE, include=FALSE}
## if running on Desktop computer, update location of Binaries
# fish_dets_20kHzPP<- updateFiles(fish_dets_20kHzPP, bin = here('../Fisher/Binaries'), db = here('../Fisher/PAM20209_Zahn_narluga_fisher.sqlite3'))
# 
# kong_dets_20kHzPP <- updateFiles(kong_dets_20kHzPP, bin = here('../Kong_Oscar/Binaries'), db = here('../Kong_Oscar/PAM20209_Zahn_narluga_kong.sqlite3'))
```

```{r function for plots, results = FALSE, include=FALSE}
## function to plot concatenated click spectrogram and mean power spectrum for each acoustic event
spec_plot <- function(Acoustic_Study) {
  ## loop through each event
  for (i in 1:length(Acoustic_Study@events)) {
    ## plot
    calculateAverageSpectra(Acoustic_Study,wl=512,evNum=Acoustic_Study@events[[i]]@id,
                            sort=FALSE,noise=FALSE,filterfrom_khz=0,
                            title=paste0("Event #", Acoustic_Study@events[[i]]@id),norm=TRUE)

  }
}
```

```{r function to make a summary table of detections}
print_summary_table <- function(species_spec_df) {
  ## round numbers to two decimal places
  species_spec_df_round <- species_spec_df %>% mutate_if(is.double, round, 2)
  ## build table
  knitr::kable(species_spec_df_round, row.names=TRUE, align = c('l','r','r','r'),
               col.names = c("Event ID",
                             "# detections",
                             "16 and 25 kHz SPL difference (dB)",
                             "25 and 40 kHz SPL difference (dB)"),
               caption = "Summary of the number of detections in each acoustic event with unique ID's and the mean sound pressure level (SPL) difference between the (1) 16 and 25 kHz, and (2) 25 and 40 kHz third-octave bands.")
}

```

```{r remove noise events, results = FALSE, include=FALSE}
## remove events with little to no clicks in event
# Fish_filtered <- filter_noise_acoustic_study(Fisher19_dets_20kHzPG_15kHzPP)
Fish_filtered <- fish_dets_0kHzPP_filter %>% filter(species!='NOISE')
Kong_filtered <- kong_dets_0kHzPP_filter %>% filter(species!='NOISE')

## load csv with SPL data
fish_spl_0kHz_df <- read.csv(here('R/spreadsheets/SPL_difference/model_input/testing_fish_data_spl_diff_0kHz_highpass.csv'))
kong_spl_0kHz_df <- read.csv(here('R/spreadsheets/SPL_difference/model_input/testing_kong_data_spl_diff_0kHz_highpass.csv'))
```

# Fisher Islands

## Beluga acoustic events

### Spectrograms and mean spectra

```{r plot fisher beluga, fig.height=3.5, fig.width=9}
## filter data for beluga events
fish_beluga_dets <- Fish_filtered %>% dplyr::filter(species=='X045')

## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
spec_plot(fish_beluga_dets)
```

### Summary table

```{r summary table fisher beluga, fig.height=3.5, fig.width=9}
## print summary table
## select beluga events
fish_spl_beluga_df <- fish_spl_0kHz_df %>% filter(species=='X045') %>% 
  select(c(eventId,n_dets,spl_diff_16to25_TOL_dB,spl_diff_25to40_TOL_dB))

print_summary_table(fish_spl_beluga_df)

```

\vspace{0.5in}

## Narwhal acoustic events

### Spectrograms and mean spectra

```{r plot fisher narwhal, fig.height=3.5, fig.width=9}
## filter data for narwhal events
fish_narwhal_dets <- Fish_filtered %>% dplyr::filter(species=='X085')

## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
spec_plot(fish_narwhal_dets)
```

### Summary table

```{r summary table fisher narwhal, fig.height=3.5, fig.width=9}
## print summary table
## select narwhal events
fish_spl_narwhal_df <- fish_spl_0kHz_df %>% filter(species=='X085') %>%
  select(c(eventId,n_dets,spl_diff_16to25_TOL_dB,spl_diff_25to40_TOL_dB))

print_summary_table(fish_spl_narwhal_df)

```


\vspace{0.5in}

# Kong Oscar

## Beluga acoustic events

### Spectrograms and mean spectra

```{r plot kong beluga, fig.height=3.5, fig.width=9}
## filter data for beluga events
kong_beluga_dets <- Kong_filtered %>% dplyr::filter(species=='X045')

## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
spec_plot(kong_beluga_dets)
```

### Summary table

```{r summary table kong beluga, fig.height=3.5, fig.width=9}
## print summary table
## select beluga events
kong_spl_beluga_df <- kong_spl_0kHz_df %>% filter(species=='X045') %>%
  select(c(eventId,n_dets,spl_diff_16to25_TOL_dB,spl_diff_25to40_TOL_dB))

print_summary_table(kong_spl_beluga_df)

```

\vspace{0.5in}

## Narwhal acoustic events

### Spectrograms and mean spectra

```{r plot kong narwhal, fig.height=3.5, fig.width=9}
## filter data for narwhal events
kong_narwhal_dets <- Kong_filtered %>% dplyr::filter(species=='X085')

## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
spec_plot(kong_narwhal_dets)
```

### Summary table

```{r summary table kong narwhal, fig.height=3.5, fig.width=9}
## print summary table
## select narwhal events
kong_spl_narwhal_df <- kong_spl_0kHz_df %>% filter(species=='X085') %>%
  select(c(eventId,n_dets,spl_diff_16to25_TOL_dB,spl_diff_25to40_TOL_dB))

print_summary_table(kong_spl_narwhal_df)

```

\vspace{0.5in}
