---
title: "Appendix A"
subtitle: "Unlabeled Acoustic Event Data Summary"
author: "M. J. Zahn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
geometry:
- top=0.75in
- bottom=0.75in
- left=0.5in
- right=0.5in
---

\vspace{0.5in}
***
\vspace{0.5in}

**Summary**

This appendix provides an overview of the data included in all acoustic events (1 hour resolution) from the Kong Oscar and Fisher Islands moorings. Events that had fewer than 30 detections contained false positive detections (i.e., noise) and thus were removed.

A summary table is provided for each category showing the number of detections for each acoustic event and the sound pressure level (SPL) difference (in dB) between selected frequency bands.

There are two figures for each event:

- Title: Acoustic event number that provides a unique numeric ID for each 6-hour acoustic event.
- Left: Concatenated click spectrogram where darker orange colors represent higher energy content.
- Right: Mean power spectrum for all detections within the event.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load packages, results = FALSE, include=FALSE}
## load required packages
library(here)
library(ggplot2)
library(PAMpal)
library(dplyr)
```

```{r load data, results = FALSE, include=FALSE}
## load Fisher data
# load(here("R/final_test/Rdata/Fisher19_20kHzPG_15kHzPP_6hourly.rdata"))
# load(here("R/Rdata/Fish_20kHzPG_20kHzPP_6hr_wavTime.rdata"))
load(here('R/Rdata/1hr/fish_20kHzPG_20kHzPP_1hr_wavTime_unfiltered.rdata'))

## load Kong Oscar data
# load(here("R/Rdata/Kong_20kHzPG_20kHzPP_6hr_wavTime.rdata"))
load(here('R/Rdata/1hr/kong_20kHzPG_20kHzPP_1hr_wavTime_unfiltered.rdata'))

```

```{r update Binaries location on path, results = FALSE, include=FALSE}
## if running on Desktop computer, update location of Binaries
# fish_dets_20kHzPP<- updateFiles(fish_dets_20kHzPP, bin = here('../Fisher/Binaries'), db = here('../Fisher/PAM20209_Zahn_narluga_fisher.sqlite3'))
# 
# kong_dets_20kHzPP <- updateFiles(kong_dets_20kHzPP, bin = here('../Kong_Oscar/Binaries'), db = here('../Kong_Oscar/PAM20209_Zahn_narluga_kong.sqlite3'))
```


```{r function to filter out noise events, results = FALSE, include=FALSE}
filter_noise_acoustic_study <- function(Acoustic_Study){
  ## remove events that have no detections or very few (< 30 detections)
  events_list <- Acoustic_Study@events
  
  for (i in 1:length(events_list)) {
    # start counter at zero
    n_clicks <- 0
    
    # loop through detectors and get total number of clicks for each event
    for (j in 1:length(events_list[[i]]@detectors)) {
      if (length(events_list[[i]]@detectors) == 0) {
        n_clicks <- 0
      }
      if (length(events_list[[i]]@detectors) == 1) {
        n_clicks <- length(events_list[[i]]@detectors[[j]]$UID)
      }
      else if (length(events_list[[i]]@detectors) > 1){
        n_clicks_det <- length(events_list[[i]]@detectors[[j]]$UID)
        n_clicks <- n_clicks + n_clicks_det
      }
    }
    if (n_clicks<30) {
      print(paste0("dropped event #", events_list[[i]]@id))
      Acoustic_Study@events[[events_list[[i]]@id]] <- NULL
    }
  }
  return(Acoustic_Study)
}

```

```{r function for plots, results = FALSE, include=FALSE}
## function to plot concatenated click spectrogram and mean power spectrum for each acoustic event
spec_calc_and_plot <- function(Acoustic_Study) {
  # make copy of Acoustic_Study object because we will modify it
  Acoustic_Study_thinned <- Acoustic_Study
  
  # create empty data frame to fill
  species_spec_df <- data.frame(evNum=c(),band_diff_17to23_dB=c(),
                                band_diff_23to42_dB=c(),n_dets=c(),n_dets_thinned=c())
  
  for (i in 1:length(Acoustic_Study_thinned@events)) {
    ## if the event has more than 1,000 clicks and under 10,000, thin the data by 10
    ## (i.e., take every 10th observation)
    if (nClicks(Acoustic_Study_thinned@events[[i]])>1000 & nClicks(Acoustic_Study_thinned@events[[i]])<=10000) {
      Acoustic_Study_thinned@events[[i]] <- Acoustic_Study_thinned@events[[i]] %>% 
        dplyr::filter(row_number(UID) %% 10 == 0)
    }
    ## take eveyr 100th click if event has >10,000 clicks
    if (nClicks(Acoustic_Study_thinned@events[[i]])>10000) {
      Acoustic_Study_thinned@events[[i]] <- Acoustic_Study_thinned@events[[i]] %>% 
        dplyr::filter(row_number(UID) %% 100 == 0)
    }

    ## then plot
    avSpec_all <- calculateAverageSpectra(Acoustic_Study_thinned,wl=512,evNum=Acoustic_Study_thinned@events[[i]]@id,
                                          sort=FALSE,noise=FALSE,filterfrom_khz=0,
                                          title=paste0("Event #", Acoustic_Study_thinned@events[[i]]@id))
    
    ## extract frequency values and mean spectra
    freq <- avSpec_all$freq
    avgSpec <- avSpec_all$avgSpec
  
    ## extract spectral data for specified TOL band
    ## 173 is our system clipping level
    band_15to19kHz_dB <- (avgSpec[which(freq>15000 & freq<19000)]+173) # 15-19 kHz TOL
    band_21to25kHz_dB <- (avgSpec[which(freq>21000 & freq<25000)]+173) # 21-25 kHz TOL
    band_40to44kHz_dB <- (avgSpec[which(freq>40000 & freq<44000)]+173) # 40-44 kHz TOL
  
    ## because dB is on a log scale, we must convert to linear scale and then take the mean
    ## after taking the mean, we can convert back to log scale
    band_15to19kHz_mean_dB <- 20*log10(mean(10^(band_15to19kHz_dB/20))) # 15-19 kHz TOL
    band_21to25kHz_mean_dB <- 20*log10(mean(10^(band_21to25kHz_dB/20))) # 21-25 kHz TOL
    band_40to44kHz_mean_dB <- 20*log10(mean(10^(band_40to44kHz_dB/20))) # 40-44 kHz TOL
  
    ## calculate difference between freq bands and convert back to dB
    band_diff_17to23_dB <- band_21to25kHz_mean_dB - band_15to19kHz_mean_dB
    band_diff_23to42_dB <- band_40to44kHz_mean_dB - band_21to25kHz_mean_dB
  
    ## calculate how many unique clicks are in the dataset
    n_clicks <- nClicks(Acoustic_Study@events[[Acoustic_Study@events[[i]]@id]], distinct=TRUE) 
    n_clicks_thin <- nClicks(Acoustic_Study_thinned@events[[Acoustic_Study_thinned@events[[i]]@id]], distinct=TRUE) 
  
    ## save mean spectrum
    df_avgSpec <- data.frame(t(avgSpec))
    
    ## build dataframe with SPL difference and click number data
    df_SPL_diff <- data.frame(evNum=Acoustic_Study_thinned@events[[i]]@id,
                              band_diff_17to23_dB=band_diff_17to23_dB, 
                              band_diff_23to42_dB=band_diff_23to42_dB,
                              n_dets=n_clicks,
                              n_dets_thinned=n_clicks_thin)
  
    ## join both dataframes
    event_spec_df <- cbind(df_SPL_diff, df_avgSpec)
    
    # append dataframe
    species_spec_df <- rbind(species_spec_df, event_spec_df)
  }
  return(species_spec_df)
}

```

```{r function to make a summary table of detections}
print_summary_table <- function(species_spec_df) {
  ## select columns we want in final table
  species_spec_df_select <- species_spec_df %>% select(evNum:n_dets_thinned)
  ## round numbers to two decimal places
  species_spec_df_select <- species_spec_df_select %>% mutate_if(is.double, round, 2)
  ## build table
  knitr::kable(species_spec_df_select, row.names=TRUE, align = c('l','r','r','r','r'),
               col.names = c("Event #", 
                             "15-19 kHz and 21-25 kHz SPL difference (dB)",
                             "21-25 kHz and 40-44 kHz SPL difference (dB)",
                             "# detections",
                             "# thinned detections"),
               caption = "Summary of the number of detections in each acoustic event and the mean sound pressure level (SPL) difference (dB) between the 15-19, 21-25, and 40-44 kHz frequency bands. Number of thinned detections provides the number of detections used in the SPL difference calulcations for large acoustic events (i.e., >1,000 detections).")
}

```


```{r run filtering function, results = FALSE, include=FALSE}
## remove events with little to no clicks in event
# Fish_filtered <- filter_noise_acoustic_study(Fisher19_dets_20kHzPG_15kHzPP)
Fish_filtered <- filter_noise_acoustic_study(fish_dets_20kHzPP)
Kong_filtered <- filter_noise_acoustic_study(kong_dets_20kHzPP)
```

# Fisher Islands

## Fisher Islands acoustic events

```{r plot fisher, fig.height=3.5, fig.width=9}
## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
fish_dets_spec_df <- spec_calc_and_plot(Fish_filtered)

## save dataframe as csv
write.csv(fish_dets_spec_df, here('R/spreadsheets/SPL_difference/fish_spectral_data_1hr.csv'))

## print summary table
print_summary_table(fish_dets_spec_df)

```

\vspace{0.5in}

# Kong Oscar

## Kong Oscar acoustic events

```{r plot kong, fig.height=3.5, fig.width=9}
## configure settings to plot mean spectrum and spectrogram side by side
par(mfrow=c(1,2))

## plot concatenated click spectrogram and mean power spectra
kong_dets_spec_df <- spec_calc_and_plot(Kong_filtered)

## save dataframe as csv
write.csv(kong_dets_spec_df, here('R/spreadsheets/SPL_difference/kong_spectral_data_1hr.csv'))

## print summary table
print_summary_table(kong_dets_spec_df)

```

\vspace{0.5in}
